School Socioeconomic Status Context and Social Skills in Children

### 2020-10-29 manuscript.Rmd
Addition of all authors

- name: Philip Parker
  affiliation: '1'
  corresponding: yes
  address: 33 Berry St, North Sydney, NSW, 2060, Australia
  email: philip.parker@acu.edu.au
  role:
  - Conceptualization
  - Formal Analysis
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
- name: Taren Sanders
  affiliation: '1'
  # corresponding: no
  # address: 33 Berry St, North Sydney, NSW, 2060, Australia
  # email: taren.sanders@acu.edu.au
  role:
  - Conceptualization
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
- name: Jake Anders
  affiliation: '2'
  # corresponding: no
  # address: 33 Berry St, North Sydney, NSW, 2060, Australia
  # email: taren.sanders@acu.edu.au
  role:
  - Conceptualization
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
- name: Michael Noetel
  affiliation: '1,3'
  # corresponding: no
  # address: 33 Berry St, North Sydney, NSW, 2060, Australia
  # email: taren.sanders@acu.edu.au
  role:
  - Conceptualization
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
shorttitle: School SES and Social Skills
output:
  pdf: papaja::apa6_pdf
  html_document:
    fig_caption: yes
    toc: yes
  pdf_document:
    toc: yes
  word: papaja::apa6_doc
  html_notebook:
    fig_caption: yes
    toc: yes
    highlight: haddock
    theme: spacelab
authornote: |
abstract: |2
  As researchers and policymakers try to understand the consequences of SES for educational and occupational attainment, they increasingly consider social skills to be critical. Yet little research has considered how the school socio-economic context may differentially promote or hinder social skill development. In a representative longitudinal sample of  `r nrow(data)` Australian 8-year olds, we tested the association between school average socioeconomic status and social skills as reported by parents and teachers. All models controlled for prior social skills at age 4, and additional school and student covariates. We found that school context was associated with social skill assimilation: controlling for individual socioeconomic status, children in more advantaged schools had more prosocial behavior and fewer peer and conduct problems. We found a consistent interaction between individual and school average socioeconomic status that suggested assimilation effects were only present for children from low socioeconomic backgrounds. We found that children from low socioeconomic backgrounds enter school with lower average social skills. They are more likely to be enrolled in more disadvantaged schools. Finally, the social skills of children from low socioeconomic backgrounds worsen in low SES contexts. Taking the evidence together, social stratified schools may harm children from low SES backgrounds while not benefiting children from high SES backgrounds.
keywords: social skills; assimilation effects; socioeconomic status; school context
wordcount: X
bibliography: r-references.bib
floatsintext: no
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
documentclass: apa6
classoption: man
affiliation:
- id: '1'
  institution: Institute for Positive Psychology and Education, Australian Catholic
    University
- id: '2'
  institution: UCL Institute of Education, UCL
- id: '3'
  institution: School of Health and Behavioural Sciences, Australian Catholic University
---

>>>>

- name: Philip Parker
  affiliation: '1'
  corresponding: yes
  address: 33 Berry St, North Sydney, NSW, 2060, Australia
  email: philip.parker@acu.edu.au
  role:
  - Conceptualization
  - Formal Analysis
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
- name: Taren Sanders
  affiliation: '1'
  # corresponding: no
  # address: 33 Berry St, North Sydney, NSW, 2060, Australia
  # email: taren.sanders@acu.edu.au
  role:
  - Conceptualization
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
- name: Jake Anders
  affiliation: '2'
  role:
  - Conceptualization
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
- name: Nikki Shure
  affiliation: '2'
  role:
  - Conceptualization
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
- name: John Jerrim
  affiliation: '2'
  role:
  - Conceptualization
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
- name: Michael Noetel
  affiliation: '1,3'
  role:
  - Conceptualization
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
- name: Joseph Ciarrochi
  affiliation: '1'
  role:
  - Conceptualization
  - Writing - Original Draft Preparation
  - Writing - Review & Editing
shorttitle: School SES and Social Skills
output:
  pdf: papaja::apa6_pdf
  html_document:
    fig_caption: yes
    toc: yes
  pdf_document:
    toc: yes
  word: papaja::apa6_doc
  html_notebook:
    fig_caption: yes
    toc: yes
    highlight: haddock
    theme: spacelab
authornote: |
abstract: |2
  As researchers and policymakers try to understand the consequences of SES for educational and occupational attainment, they increasingly consider social skills to be critical. Yet little research has considered how the school socio-economic context may differentially promote or hinder social skill development. In a representative longitudinal sample of  `r nrow(data)` Australian 8-year olds, we tested the association between school average socioeconomic status and social skills as reported by parents and teachers. All models controlled for prior social skills at age 4, and additional school and student covariates. We found that school context was associated with social skill assimilation: controlling for individual socioeconomic status, children in more advantaged schools had more prosocial behavior and fewer peer and conduct problems. We found a consistent interaction between individual and school average socioeconomic status that suggested assimilation effects were only present for children from low socioeconomic backgrounds. We found that children from low socioeconomic backgrounds enter school with lower average social skills. They are more likely to be enrolled in more disadvantaged schools. Finally, the social skills of children from low socioeconomic backgrounds worsen in low SES contexts. Taking the evidence together, social stratified schools may harm children from low SES backgrounds while not benefiting children from high SES backgrounds.
keywords: social skills; assimilation effects; socioeconomic status; school context
wordcount: X
bibliography: r-references.bib
floatsintext: no
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
documentclass: apa6
classoption: man
affiliation:
- id: '1'
  institution: Institute for Positive Psychology and Education, Australian Catholic
    University
- id: '2'
  institution: UCL Institute of Education, UCL
- id: '3'
  institution: School of Health and Behavioural Sciences, Australian Catholic University
---

### 2020-10-29 manuscript.tex
Addition of all authors

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={School Socioeconomic Status Context and Social Skills in Children},
  pdfauthor={Philip Parker1, Taren Sanders1, Jake Anders2, \& Michael Noetel1,3},
  pdflang={en-EN},
  pdfkeywords={social skills; assimilation effects; socioeconomic status; school context},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{School Socioeconomic Status Context and Social Skills in Children}
\author{Philip Parker\textsuperscript{1}, Taren Sanders\textsuperscript{1}, Jake Anders\textsuperscript{2}, \& Michael Noetel\textsuperscript{1,3}}
\date{}

@@

The authors made the following contributions. Philip Parker: Conceptualization, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Taren Sanders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Jake Anders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Michael Noetel: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={School Socioeconomic Status Context and Social Skills in Children},
  pdfauthor={Philip Parker1, Taren Sanders1, Jake Anders2, Nikki Shure2, John Jerrim2, Michael Noetel1,3, \& Joseph Ciarrochi1},
  pdflang={en-EN},
  pdfkeywords={social skills; assimilation effects; socioeconomic status; school context},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{School Socioeconomic Status Context and Social Skills in Children}
\author{Philip Parker\textsuperscript{1}, Taren Sanders\textsuperscript{1}, Jake Anders\textsuperscript{2}, Nikki Shure\textsuperscript{2}, John Jerrim\textsuperscript{2}, Michael Noetel\textsuperscript{1,3}, \& Joseph Ciarrochi\textsuperscript{1}}
\date{}

@@

The authors made the following contributions. Philip Parker: Conceptualization, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Taren Sanders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Jake Anders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Nikki Shure: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; John Jerrim: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Michael Noetel: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Joseph Ciarrochi: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.
